%% The following is a directive for TeXShop to indicate the main file
%%!TEX root = diss.tex

\newcommand{\numstarted}{90}
\newcommand{\numfinished}{65}

\chapter{Survey}
\label{ch:Survey}

\noindent We deployed a survey to investigate the type and frequency of
reachability questions that developers may encounter during their day-to-day 
activities.

\section{Method}
\label{sec:method}

\noindent The survey was developed with the goal of maximizing the relevance of 
the presented questions to actual scenarios that developers would encounter in
their day-to-day work.
We drew upon our previous work as software developers as well as the literature
for reachability questions \cite{latoza-2010-reach} and hard-to-answer
questions about code that developers ask \cite{latoza-2010-hard-questions}.
From these sources, we drafted a set of 10 questions that we hypothesized 
developers may ask as they performed change tasks or explored their programs.

\par After piloting the survey in a laboratory setting with 7 graduate
students, we finalized a survey with 9 questions with corresponding code
excerpts with a time-to-completion of about 10 minutes.
The survey was deployed entirely online, and consisted of:
\begin{itemize}
    \item 5 demographic questions.
    \item 9 Likert-scale questions that each represented a hypothetical 
          scenario associated with a program comprehension or reachability 
          question.
\end{itemize}

% TODO: describe the survey (use tables?) attach it to this. Make a footnote,

\subsection{Participant Recruitment}
\label{subsec:ParticipantRecruitment}

\noindent
We distributed the survey on Twitter using our personal accounts.
In addition to Twitter, the survey was distributed using mailing lists to
professional software developers.

% TODO: come back to this once we can actually say more about  what we  build
\par As our tool in its current state is targeted toward developers who
use statically-typed programming languages, we excluded respondents who have
never used a statically-typed language as their primary development language.
We also excluded individuals who do not write code at least once a month,
as they may not be the primary audience for an eventual tool.

\section{Survey Results}
\label{sec:SurveyResults}

The survey was started \numstarted{} times, and finished \numfinished{} times
in total. 
This yielded a survey completion rate of \printpcnt{\numfinished}{\numstarted}.

\section{Threats to Validity}
\label{sec:ThreatsToValidity}

% Previously: External Validity

\noindent One of the ways which we distributed our survey was via Twitter.  
This means that part of our results may be specific to the responses of
software developers who use Twitter.
Consequently, it may be challenging to fully generalize our results across
the general population of developers who may not use Twitter.
In an attempt to minimize this threat to validity, we also distributed the 
survey using mailing lists to professional software developers, under the
assumption that most professional software developers use email
\cite{gousios-2016-work-practices}.

% Previously: Construct Validity

\par To create a set of survey questions that would closely represent the
questions and scenarios that developers reported to be most challenging
\cite{latoza-2010-hard-questions, latoza-2010-reach}, we designed a small Java
project represening an e-commerce system.
The Java methods presented in each survey question may not be entirely
representative of the actual code and systems that the respondents may have
experience with.
We deemed this to be an acceptable threat, as the code excerpts we drew from
the sample project were writen to distill the core concepts behind
reachability and program comprehension questions into a succinct representation.
Having overly-complicated code excerpts may have impacted the study by
increasing time-to-completion and drop rate.

\par The code excerpts in our survey were static elements, meaning that the
only way our participants could interact with the code was to read them
in isolation.
Although this furthered our design goal of making the questions encompass
the core concepts behind reachability questions, it did not accurately
capture how developers interact with code (e.g., IDE-based navigation, 
other workflows and tooling) in program comprehension tasks.
To mitigate this threat, we may have had to design our survey around an IDE 
that our participants would be able to use.
However, it is likely that this would have introduced another threat to
validity; our results may be affected by the proficiency of our participants
with the IDE we select to use in our study.
Consequently, we deemed the lack of interactable code excerpts to be an 
acceptable threat.

\par Another threat to validity is the fact that all the code
excerpts included in our survey were written in Java.
This does not account for the fact that our participants may have varying
levels of proficiency with Java.
Consequently, we may be conflating our results with the proficiency of our
participants with Java.
That being said, we needed to select a language for our study that would
have a high likelihood of being familiar to our target audience.
Java's popularity as a development language 
\cite{so-2021-dev-survey, jetbrains-2021-dev-survey} meant that it was a 
suitable choice.

\endinput

